Missing data: a simple introduction

### 2020-03-15 Draft1.tex
Minor changes

	\subsection{Congeniality: when the imputer's and analysts models differ}
	As mentioned in the introduction, modern data analysis' multiple phases mean that the assumptions underlying each phase and the relationship between the phases is an important object of study. Differences might cause the subsequent analysis to be biased, or in the worst case, invalid. Meng's [meng 1994] concept of congeniality provides a theoretical framework from which to evaluate these differences and most of this section is derived from his paper.

@@

	Congeniality is an important concept in imputation and has a big impact on the statistical validity of MI. In short, uncongeniality arises when the analysis procedure differs substantially from the imputation model. There are many reasons for such difference, for example, the imputer might wish to include as many variables as possible to achieve an accurate imputation, while the analyst might be interested in only a subset of the covariates and the relationship between these. However, as Meng [1994] states, ''If the imputer's model is reasonably accu-
	rate, then following the multiple-imputation recipe
	prevents the analyst from producing inferences with
	serious nonresponse biases''. Meng goes on to show that even with an uncongenial imputation model, inferences performed on a MI dataset still yield better inference than the best possible incomplete data inference.

@@

	\textbf{Last Observation Carried Forward (LOCF):x}
	This method is mostly used in longitudinal studies where individuals drop out, withdraw, or are lost over the course of the studies. LOCF involves imputing the missing values for an individual with the last observed value for that individual. Leaving aside questions of precision, LOCF produces biased estimators in all except very specific MNAR circumstances (Kenward and Molenberghs 2009 get article). With LOCF we are effectively assuming that for each individual that drops out their responses do not change at all for the remainder of the study, a highly unrealistic assumption. It can be concluded, therefore, that LOCF is the worst imputation method of those mentioned here, and all the other methods are preferable. According [Missing data] "LOCV is neither valid under general assumptions nor based on statistical principles, it is not a sensible method, and should not be used".

@@

	Xia and Meng (2017) state that ''When the imputation model class and the (embedded) analyst’s model class
	differ, the behavior of Rubin’s rules becomes very complicated, capable of producing inconsistent variance estimators, a matter that has received recurrent
	criticisms''. They argue that in the uncongenial case , which is often the rule rather than the exception, doubling Rubin's variance measure will never underestimate the true variance. This holds for any degree of uncongeniality, and provides a surprisingly simple solution to the problem of uncongeniality. For Univariate data, Xia and Meng (2017) suggest replacing $T$ with $\tilde{T}= (\sqrt{\bar{Var}}+\sqrt{B})^{2}+\frac{1}{M}B_$. These methods may be viewed as highly conservative, but in the context of MI where uncongeniality between analysts and imputer is unknown and potentially large, it pays to be as insured as possible against invalid inference. However, even if these methods are not used and the models are uncongenial, we note that Meng (1994) has shown that MI is still superior to incomplete data procedures.

>>>>

	\subsection{Congeniality: when the imputer's and analysts models differ}
	Congeniality is an important concept in imputation and has a big impact on the statistical validity of MI. In short, uncongeniality arises when the analysis model differs substantially from the imputation model. There are many reasons for such difference, for example, the imputer might wish to include as many variables as possible to achieve an accurate imputation, while the analyst might be interested in only a subset of the covariates and the relationship between these. However, as Meng [1994] states, ''If the imputer's model is reasonably accu-
	rate, then following the multiple-imputation recipe
	prevents the analyst from producing inferences with
	serious nonresponse biases''. Meng goes on to show that even with an uncongenial imputation model, inferences performed on a MI dataset still yield better inference than the best possible incomplete data inference.

	As mentioned in the introduction, modern data analysis' multiple phases mean that the assumptions underlying each phase and the relationship between the phases is an important object of study. Differences might cause the subsequent analysis to be biased, or in the worst case, invalid. Meng's [meng 1994] concept of congeniality provides a theoretical framework from which to evaluate these differences and most of this section is derived from his paper.

@@

@@

	\textbf{Last Observation Carried Forward (LOCF):}
	This method is mostly used in longitudinal studies where individuals drop out, withdraw, or are lost over the course of the studies. LOCF involves imputing the missing values for an individual with the last observed value for that individual. Leaving aside questions of precision, LOCF produces biased estimators in all except very specific MNAR circumstances (Kenward and Molenberghs 2009 get article). With LOCF we are effectively assuming that for each individual that drops out their responses do not change at all for the remainder of the study, a highly unrealistic assumption. It can be concluded, therefore, that LOCF is the worst imputation method of those mentioned here, and all the other methods are preferable. According [Missing data] "LOCV is neither valid under general assumptions nor based on statistical principles, it is not a sensible method, and should not be used".

@@

	Xia and Meng (2017) state that ''When the imputation model class and the (embedded) analyst’s model class
	differ, the behavior of Rubin’s rules becomes very complicated, capable of producing inconsistent variance estimators, a matter that has received recurrent
	criticisms''. They argue that in the uncongenial case , which is often the rule rather than the exception, doubling Rubin's variance measure will never underestimate the true variance. This holds for any degree of uncongeniality, and provides a surprisingly simple solution to the problem of uncongeniality. For Univariate data, Xia and Meng (2017) suggest replacing $T$ with $\tilde{T}= (\sqrt{\bar{Var}}+\sqrt{B})^{2}+\frac{1}{M}B$. These methods may be viewed as highly conservative, but in the context of MI where uncongeniality between analysts and imputer is unknown and potentially large, it pays to be as insured as possible against invalid inference. However, even if these methods are not used and the models are uncongenial, we note that Meng (1994) has shown that MI is still superior to incomplete data procedures.
